%!TEX root = ../thesis.tex
\chapter{Some Appendix Chapter}

\section{Hamburg Street Descriptions}

\textbf{Agathe-Lasch-Weg} \par
Agathe-Lasch-Weg begins at the intersection with Holmbrook, Paul-Ehrlich-Straße. \par
Emkendorfstraße then branches off to the right. \par
Roosens Weg then branches off to the right. \par
Halbmondsweg then branches off to the left. \par
Klein Flottbeker Weg then branches off to the right. \par
Reventlowstraße then branches off to the right. \par

\bigskip % Add vertical space between street entries

\textbf{Ansorgestraße} \par
Ansorgestraße begins at the intersection with Emkendorfstraße, Liebermannstraße. \par
Roosens Weg then branches off to the right. \par

\bigskip

\textbf{Behringstraße} \par
Behringstraße begins at the intersection with ['Walderseestraße', 'Behringstraße']. \par

\bigskip

\textbf{Bernadottestraße} \par
Bernadottestraße begins at the intersection with Droysenstraße. \par
Halbmondsweg then branches off to the right. \par
Halbmondsweg then branches off to the right. \par
Poppes Weg then branches off to the right. \par
Roosens Weg then branches off to the left. \par
Schlagbaumtwiete then branches off to the right. \par
Elbblöcken then branches off to the right. \par

\bigskip

\textbf{Borchlingweg} \par
Borchlingweg begins at the intersection with Ansorgestraße, Halbmondsweg. \par
Stindeweg then branches off to the right. \par
Langmaackweg then branches off to the right. \par

\bigskip

\textbf{Corinthstraße} \par
Corinthstraße begins at the intersection with Bernadottestraße. \par
Elbblöcken then branches off to the right. \par
Elbchaussee then branches off to the right. \par

\bigskip

\textbf{Droysenstraße} \par
Droysenstraße begins at the intersection with Jungmannstraße. \par
Preußerstraße then branches off to the left. \par

\bigskip

\textbf{Elbblöcken} \par
Elbblöcken begins at the intersection with Bernadottestraße. \par
Leipoldstieg then branches off to the right. \par

\bigskip

\textbf{Elbchaussee} \par
Elbchaussee begins at the intersection with Schlagbaumtwiete. \par
Corinthstraße then branches off to the left. \par

\bigskip

\textbf{Emkendorfstraße} \par
Emkendorfstraße begins at the intersection with Agathe-Lasch-Weg. \par
Röpers Weide then branches off to the left. \par
Lavaterweg then branches off to the left. \par
Reventlowstraße then branches off to the right. \par

\bigskip

\textbf{Ernst-August-Straße} \par
Ernst-August-Straße begins at the intersection with Liebermannstraße. \par
Roosens Weg then branches off to the left. \par

\bigskip

\textbf{Halbmondsweg} \par
Halbmondsweg begins at the intersection with Bernadottestraße. \par
Bernadottestraße then branches off to the left. \par
Ansorgestraße then branches off to the left. \par
Borchlingweg then branches off to the right. \par
Agathe-Lasch-Weg then branches off to the right. \par
Klein Flottbeker Weg then branches off to the left. \par
Reventlowstraße then branches off to the right. \par

\bigskip

\textbf{Hammerichstraße} \par
Hammerichstraße begins at the intersection with Jungmannstraße. \par
Walderseestraße then branches off to the left. \par

\bigskip

\textbf{Holmbrook} \par
Holmbrook begins at the intersection with Agathe-Lasch-Weg, Paul-Ehrlich-Straße. \par

\bigskip

\textbf{Jungmannstraße} \par
Jungmannstraße begins at the intersection with Reventlowstraße. \par
Preußerstraße then branches off to the left. \par
Droysenstraße then branches off to the left. \par

\bigskip

\textbf{Klein Flottbeker Weg} \par
Klein Flottbeker Weg begins at the intersection with Agathe-Lasch-Weg, Halbmondsweg, Reventlowstraße. \par
Zickzackweg then branches off to the left. \par
Droysenstraße then branches off to the left. \par

\bigskip

\textbf{Langmaackweg} \par
Langmaackweg begins at the intersection with Borchlingweg. \par

\bigskip

\textbf{Lavaterweg} \par
Lavaterweg begins at the intersection with Emkendorfstraße. \par

\bigskip

\textbf{Leipoldstieg} \par
Leipoldstieg begins at the intersection with Elbblöcken. \par

\bigskip

\textbf{Liebermannstraße} \par
Liebermannstraße begins at the intersection with Ansorgestraße, Emkendorfstraße. \par
Ernst-August-Straße then branches off to the right. \par
Bernadottestraße then branches off to the right. \par

\bigskip

\textbf{Lobsienweg} \par
Lobsienweg begins at the intersection with Droysenstraße. \par

\bigskip

\textbf{Meistertwiete} \par
Meistertwiete begins at the intersection with Bernadottestraße. \par

\bigskip

\textbf{Olshausenstraße} \par
Olshausenstraße begins at the intersection with Reventlowstraße. \par
Reventlowstraße then branches off to the right. \par
Walderseestraße then branches off to the left. \par

\bigskip

\textbf{Paul-Ehrlich-Straße} \par
Paul-Ehrlich-Straße begins at the intersection with Agathe-Lasch-Weg, Holmbrook. \par

\bigskip

\textbf{Poppes Weg} \par
Poppes Weg begins at the intersection with Bernadottestraße. \par
Ansorgestraße then branches off to the left. \par

\bigskip

\textbf{Preußerstraße} \par
Preußerstraße begins at the intersection with Jungmannstraße. \par
Droysenstraße then branches off to the right. \par

\bigskip

\textbf{Reventlowstraße} \par
Reventlowstraße begins at the intersection with Olshausenstraße. \par
Walderseestraße then branches off to the left. \par

\bigskip

\textbf{Roosens Park} \par
Roosens Park begins at the intersection with Ansorgestraße. \par

\bigskip

\textbf{Roosens Weg} \par
Roosens Weg begins at the intersection with Ansorgestraße. \par
Agathe-Lasch-Weg then branches off to the left. \par
Reventlowstraße then branches off to the right. \par

\bigskip

\textbf{Röpers Weide} \par
Röpers Weide begins at the intersection with Walderseestraße. \par
Emkendorfstraße then branches off to the right. \par

\bigskip

\textbf{Schlagbaumtwiete} \par
Schlagbaumtwiete begins at the intersection with Elbchaussee. \par
Bernadottestraße then branches off to the right. \par

\bigskip

\textbf{Slevogtstieg} \par
Slevogtstieg begins at the intersection with Elbblöcken. \par

\bigskip

\textbf{Stindeweg} \par
Stindeweg begins at the intersection with Borchlingweg. \par

\bigskip

\textbf{Taxusweg} \par
Taxusweg begins at the intersection with Halbmondsweg. \par

\bigskip

\textbf{Walderseestraße} \par
Walderseestraße begins at the intersection with Reventlowstraße. \par
Reventlowstraße then branches off to the right. \par

\bigskip

\textbf{Zickzackweg} \par
Zickzackweg begins at the intersection with Klein Flottbeker Weg. \par

\bigskip

\textbf{Zypressenweg} \par
Zypressenweg begins at the intersection with Halbmondsweg. \par

\bigskip

\textbf{['Walderseestraße', 'Behringstraße']} \par
['Walderseestraße', 'Behringstraße'] begins at the intersection with Behringstraße. \par
Walderseestraße then branches off to the right. \par

\bigskip

\textbf{Övelgönner Hohlweg} \par
Övelgönner Hohlweg begins at the intersection with Elbchaussee, Halbmondsweg. \par